This is R markdown for project 1 in course Reproducbile research.

#Auhor : Saeedeh Salimianrizi

##preparing data
```{r preparing data, echo = TRUE}
data <- read.csv("activity.csv")
head(data)
summary(data)
```


##total number of steps taken per day
```{r total number of steps taken per day, echo = TRUE}
steps_by_day <- aggregate(steps ~ date, data, FUN = sum)
```

##histogram of the total number of steps taken each day
```{r hist, echo = TRUE}
hist(steps_by_day$steps, breaks = 10, col = "blue", main = "Total Steps Per Day", xlab = "Number of Steps")
```

##mean and median of the total number of steps taken per day
```{r mean and median, echo = TRUE}
mean(steps_by_day$steps , na.rm= TRUE)
median(steps_by_day$steps , na.rm= TRUE)
```

## time series plot of the 5-minute interval and and average number of steps taken
```{r time series plot, echo = TRUE}
average_steps_by_interval <- aggregate(steps ~ interval, data, FUN = mean)
plot(average_steps_by_interval$interval, average_steps_by_interval$steps, type = "l", main = "Average Steps~Intervals",xlab = "Intervals", ylab = "Average Steps")
```


##5-minute interval with the max number of steps
```{r 5 minute interval the max number of steps, echo =  TRUE}
max_average_step <- max(average_steps_by_interval$steps)
average_steps_by_interval[average_steps_by_interval$steps == max_average_step,]
```


##total number of missing values in the data set
```{r number of missing values in the data set, echo = TRUE}
sum(is.na(data))
```


##impute data, replacing NA values with the mean values of the same interval
```{r impute data, echo = TRUE}
data_noNA <- transform(data, steps = ifelse(is.na(data$steps), average_steps_by_interval$steps[match(data$interval, average_steps_by_interval$interval)], data$steps))
steps_by_day_noNA <- aggregate(steps ~ date, data_noNA, FUN = sum)
```


##hist of real data
```{r comparing hist of real and imputed data}
hist(steps_by_day$steps,  breaks = 10, col = "blue", main = "Total Steps Per Day", xlab = "Number of Steps")
```


## hist of imputed data
```{r host of imputed data}
hist(steps_by_day_noNA$steps, breaks = 10,  col = "red", main = "Total Steps Per Day", xlab = "Number of Steps")
```

### as you can see the histogram of real and imputed data are slightly different

##mean and the median for the total number of steps taken per day using imputed data
```{r mean anf the median for imputed data, echo = TRUE}
mean(steps_by_day_noNA$steps)
median(steps_by_day_noNA$steps)
```

### as you can see the mean value stayed the same after imputing data but median increased slightly.
##identify weekday and weekend related data
```{r weekday/weekend related data, echo = TRUE}
data_noNA$weekday <- ifelse(weekdays(as.Date(data_noNA$date))%in%c("Sunday", "Saturday"), "Weekend", "Workday")
```


## time series plot of the 5-minute interval  and the average number of steps takeN for weekday days or weekend days
```{r time series}
average_steps_by_interval_noNA <- aggregate(steps ~ interval + weekday, data_noNA, FUN = mean)

library(lattice)
xyplot(average_steps_by_interval_noNA$steps ~ average_steps_by_interval_noNA$interval|average_steps_by_interval_noNA$weekday, type = "l", 
layout = c(1,2), xlab = "Interval", ylab= "Average Steps Per Day", main = "Average Steps per Day by Interval")
```
